Exact Symbolic Dynamic Programming for Continuous POMDPs

### 2012-05-31 cpomdp/camdp-xadd/src/doc/NIPS/draft3-10page.tex
With additional pics for relevant partitions

\Begin
{
	$z:=0$;\	
		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
		{
    	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $;\
		\For {all $j$ in $\alpha_j(xd_s,xd_o)$}    
		{
		$\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$\}\ \;
		\For {all $j$ in $\delta_{j}(xd_o)$}    
		{
		$z := max(\delta_j(xd_o),z)$\;
    	\emph{// $\phi_{z_k}$ logical term on each partition of $z$}\\
   		$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$ \;
    \Return{$P(z)$} \;
    %do this for each belief in B
\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(z)$ }
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For $h=2$ the resulting $\alpha$-vectors after this regression step are functions of the observation and current state: 
\alpha_1^2(s,o) &= 
 (t_s<15)\wedge (t_s - 10 < t_o<t_s) &: 10 \\
(t_s>15)\wedge (t_s - 10 < t_o<t_s) &: -100  \\
\neg(t_s - 10 < t_o<t_s) &: 0
\hspace{4mm} 
\alpha_2^2(s,o) &= \begin{cases}
(t_s - 10 < t_o<t_s) &: -0.1 \\
\neg(t_s - 10 < t_o<t_s) &: 0
\nonumber
We need the $\alpha$-vectors to partition on the observation space, thus next we eliminate the current state variable. During this integration the current state variable is factored out by multiplying the belief in each of the current $\alpha$-vectors.  The resulting $\delta$-functions only depend on the observation: $\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.
For the belief $b_1: U[t_s;2,6]$ the two vectors above the resulting $\delta$-functions are defined below: 
\delta_{1}(o) &= 
 (2<t_o<6) &: 15 - 2.5*t_o \\
(-4<t_o<2) &: 10 \\
(-8<t_o<-4) &: 20 + 2.5*t_o 
\delta_{2}(o) &= \begin{cases}
 (2<t_o<6) &: -15 + 0.025*t_o \\
(-4<t_o<2) &: -0.1 \\
(-8<t_o<-4) &: -0.2 - 0.025*t_o 
\nonumber
The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$. According to continuous observation space defined in \cite{pascalPomdp}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space. In this work, numerical solutions are proposed to find the roots of any two observation dependant function. Here we use the symbolic power of the max-operator to find all the maximum boundary regions of two or more $\delta$-functions at the same time. For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximuum: 
\mathrm{max} \Bigg(\delta_{1}(o),\delta_{1}(o)\Bigg) &= 
 (2<t_o<6) &: 15 - 2.5*t_o \\
(-4<t_o<2) &: 10 \\
(-8<t_o<-4) &: 20 + 2.5*t_o 
\nonumber
The three partitions define 3 observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the three distinct observation partitions is found. If we integrate out the state and observation from the observation model and the belief state regardless of a particular $\alpha$-vector, it integrates to one: 
$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$
Thus we only need to multiply the indicators of each case partition in this formula to obtain the probability of each partition such that it sums to one: 
$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$
For our example the following holds: 
P(z_k)=
 (2<t_o<6) &: 0.2 \\
(-4<t_o<2) &: 0.6\\
(-8<t_o<-4) &:0.1999
\nonumber
Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space in Monahan's algorithm. 
%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities. In a lyx file, needed?
Next we present some results for 2-Dimensional continuous observation spaces.

>>>>

\Begin
{
	$z:=0$;\	
		\For {all $\alpha_j$ in  $\Gamma^{h-1}$}    
		{
    	$\alpha_j(xd_s,xd_o) := \int_{s'} P(xd_{o_i}|xd'_{s_i},a) * P(xd'_{s_i}| xd_{s_i},a)* \alpha_j $;\
		}  
		\For {all $j$ in $\alpha_j(xd_s,xd_o)$}    
		{
		$\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$\}\ \;
		}
		\For {all $j$ in $\delta_{j}(xd_o)$}    
		{
		$z := max(\delta_j(xd_o),z)$\;
    	}
    	\emph{// $\phi_{z_k}$ logical term on each partition of $z$}\\
   		$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$ \;
    \Return{$P(z)$} \;
    %do this for each belief in B
}
\caption{\footnotesize \texttt{GenRelObs}($\Gamma^h,a,b_i$) $\longrightarrow$ $P(z)$ }
\vspace{-1mm}
\end{algorithm}
\decmargin{.5em}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For $h=2$ the resulting $\alpha$-vectors after this regression step are functions of the observation and current state: 
{\footnotesize
\begin{align}
\alpha_1^2(s,o) &= 
\begin{cases}
 (t_s<15)\wedge (t_s - 10 < t_o<t_s) &: 10 \\
(t_s>15)\wedge (t_s - 10 < t_o<t_s) &: -100  \\
\neg(t_s - 10 < t_o<t_s) &: 0
\end{cases}
\hspace{4mm} 
\alpha_2^2(s,o) &= \begin{cases}
(t_s - 10 < t_o<t_s) &: -0.1 \\
\neg(t_s - 10 < t_o<t_s) &: 0
\end{cases}
\nonumber
\end{align}
}
We need the $\alpha$-vectors to partition on the observation space, thus next we eliminate the current state variable. During this integration the current state variable is factored out by multiplying the belief in each of the current $\alpha$-vectors.  The resulting $\delta$-functions only depend on the observation: $\delta_{j}(xd_o) := $ $\int_{xd_{s_i}} b_i * \alpha_j(xd_s,xd_o)$.
For the belief $b_1: U[t_s;2,6]$ the two vectors above the resulting $\delta$-functions are defined below: 
{\footnotesize
\begin{align}
\delta_{1}(o) &= 
\begin{cases}
 (2<t_o<6) &: 15 - 2.5*t_o \\
(-4<t_o<2) &: 10 \\
(-8<t_o<-4) &: 20 + 2.5*t_o 
\end{cases}
\hspace{5mm} 
\delta_{2}(o) &= \begin{cases}
 (2<t_o<6) &: -0.15 + 0.025*t_o \\
(-4<t_o<2) &: -0.1 \\
(-8<t_o<-4) &: -0.2 - 0.025*t_o 
\end{cases}
\nonumber
\end{align}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}[tbp!]
\vspace{-6mm}
\centering
\hspace{-17mm}
\includegraphics[width=0.41\textwidth]{pics/beliefs.pdf}
\hspace{-12mm}
\includegraphics[width=0.42\textwidth]{pics/delta1.pdf}
\hspace{-12mm}
\includegraphics[width=0.42\textwidth]{pics/delta2.pdf}
\hspace{-17mm}
\vspace{-10mm}
\caption{\footnotesize 
{\it (left)} Beliefs $b_1,b_2$ for 1D \texttt{Power Plant} example; 
{\it (middle)} Observation dependent function for $b_1$ that partition the observation space into 3 regions with different probabilities $P(z_i)$ ; 
{\it (right)} Relevant observation partitions and their probabilities for $b_2$.
}
\label{fig:timeSpace}
\vspace{-4mm}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
The observation dependent $\delta$-functions divide the observation space into regions which can yield the optimal policy according to the belief state $b_1$. According to continuous observation space defined in \cite{pascalPomdp}, for a 1-dimensional observation space with two or more functions of the observation, we need to find the optimal boundaries or partitions of the space. In this work, numerical solutions are proposed to find the roots of any two observation dependant function. Here we use the symbolic power of the max-operator to find all the maximum boundary regions of two or more $\delta$-functions at the same time. For the two $\delta$-functions above the following partitions of the observation space is derived after taking their maximuum: 
{\footnotesize
\begin{align}
\mathrm{max} \Bigg(\delta_{1}(o),\delta_{1}(o)\Bigg) &= 
\begin{cases}
 (2<t_o<6) &: 15 - 2.5*t_o \\
(-4<t_o<2) &: 10 \\
(-8<t_o<-4) &: 20 + 2.5*t_o 
\end{cases}
\nonumber
\end{align}
}
The three partitions define 3 observation regions which we can now use in a similar fashion to a discrete observation set if only the probability of each of the three distinct observation partitions is found. If we integrate out the state and observation from the observation model and the belief state regardless of a particular $\alpha$-vector, it integrates to one: 
$\int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*P(s) = 1$
Thus we only need to multiply the indicators of each case partition in this formula to obtain the probability of each partition such that it sums to one: 
$P(z_k) := \int_{xd_o}\int_{xd_s} P(xd_o|xd'_s,a)*P(xd'_s|xd_s,a)*b_i* \mathbb{I}[\phi_{z_k}] d_{xd_o} d_{xd_s}$
For our example the following holds: 
{\footnotesize
\begin{align}
P(z_k)=
\begin{cases}
 (2<t_o<6) &: 0.2 \\
(-4<t_o<2) &: 0.6\\
(-8<t_o<-4) &:0.1999
\end{cases} 
\nonumber
\end{align}
}
Hence we can now use the algorithms and methods of a discrete observation setting using the probabilities of the partitioned observation space in Monahan's algorithm. 
%Figure ~\ref{fig:obsPartition} shows the observation partitions obtained in the 1D problem instance for different horizons and their related probabilities. In a lyx file, needed?
Next we present some results for 2-Dimensional continuous observation spaces.
